---
title: "Day 1"
author: "Sneha Jacob"
date: "2024-09-26"
categories: [intro, babyname-analysis, line-graph]
---

## Introduction

I'm a designer in training and like most people in the creatives, my interests cover a broad spectrum ranging from 3D Modelling and Lettering to Website Designs. I'm now working to add data visualisation and into my skill set, **I hope I don't totally suck at it**!

## My First Piece of R-code

I'm doing this!!!!!

```{r}
#| label: setup
library (tidyverse)
library (ggformula)
library (babynames)

library(mosaic)
library(skimr)
library(knitr)
library(kableExtra)
```

#### Getting a list of all baby names in the USA from the year 1880

```{r}
babynames
```

```         
```

#### Filtering the presence of my name i.e. Sneha and creating a line graph with this data

```{r}
babynames %>% filter (name=="Sneha")

```

```{r}
babynames %>% filter(name=="Sneha") %>% gf_line(n~year)
```

#### Filtering the presence of the name "Trisha" and creating a line graph with this data

```{r}
babynames %>% filter (name=="Trisha")
```

```{r}
babynames %>% filter(name=="Trisha") %>% gf_line(n~year)
```

#### Filtering the presence of the name "Sarah" and creating a line graph with this data

```{r}
babynames %>% filter (name=="Sarah" | name=="Sara") 

```

```{r}
babynames %>% filter(name=="Sarah" | name=="Sara") %>% gf_line(n~year)

```

#### Glimpse:

```{r}
babynames %>% dplyr::glimpse()
```

```{r}
babynames_modified <- babynames %>%
  dplyr::mutate(
    sex = as_factor(sex),
  )
glimpse(babynames_modified)
```

```{r}
babynames_modified %>% mosaic::inspect()
```

```{r}
babynames_modified %>% skimr::skim()
```
