---
title: "Day 5 - Part 2"
author: "Sneha Manu Jacob"
date: "2024-10-10"
categories: [change, scatter-plot, correlation]
---

For count, both variable are quantitative

### Correlation

Using Scatter plot, we aim to find the correlation between 2 quantitative values. Correlation is commonly used in data analysis to determine whether, and how strongly, variables are related. We assume that the association is linear, that one variable increases or decreases a fixed amount for a unit increase or decrease in the other. The other technique that is often used in these circumstances is regression, which involves estimating the best straight line to summarise the association.

Strong Correlation: Values close to 1 or -1 (e.g., 0.8 or -0.8) show a strong relationship. Weak Correlation: Values close to 0 (e.g., 0.1 or -0.1) show a weak relationship.

**Pearson Correlation**: Measures the linear relationship between two continuous variables. It’s the most commonly used type.

**Line of best fit**- is a straight line that best represents the relationship between two variables in a scatter plot. The line of best fit helps summarize the relationship between two variables by providing a clear visual and mathematical description of how one variable (the independent variable) influences the other (the dependent variable).

```{r}
#| label: setup
library(tidyverse) # Tidy data processing and plotting
library(ggformula) # Formula based plots
library(mosaic) # Our go-to package
library(skimr) # Another Data inspection package
library(kableExtra) # Making good tables with data

library(GGally) # Corr plots
library(corrplot) # More corrplots
library(ggExtra) # Making Combination Plots

# devtools::install_github("rpruim/Lock5withR")
library(Lock5withR) # Datasets
library(palmerpenguins) # A famous dataset

library(easystats) # Easy Statistical Analysis and Charts
library(correlation) # Different Types of Correlations
# From the easystats collection of packages
```

## Hollywood movies data-set

```{r}
HollywoodMovies2011 -> movies
glimpse(movies)
```

-   **Movie (Factor)**:

    -   This is likely the title or identifier of the movie. Since it’s a factor, it means there are multiple unique movie titles.

-   **LeadStudio (Factor)**:

    -   The production studio that produced or distributed the movie. It could include studios like Warner Bros, Disney, Universal, etc.

-   **RottenTomatoes (Numeric)**:

    -   The movie’s rating or score on Rotten Tomatoes, a popular review aggregator. It's often expressed as a percentage, representing the proportion of positive reviews by critics.

-   **AudienceScore (Numeric)**:

    -   A score reflecting how audiences rated the movie, also typically on a percentage scale. This is often gathered from surveys or audience feedback platforms.

-   **Story (Factor)**:

    -   This could refer to the primary theme or storyline category of the movie (e.g., Hero’s Journey, Revenge, Coming of Age). It might represent narrative tropes or common plot structures.

-   **Genre (Factor)**:

    -   The type of movie based on its theme or tone (e.g., Action, Comedy, Drama, Horror). These are pre-defined categories common to film classification.

-   **TheatersOpenWeek (Numeric)**:

    -   The number of theaters where the movie was shown during its opening week. A larger number indicates a wide release, while a smaller number could suggest a limited release.

-   **BOAverageOpenWeek (Numeric)**:

    -   Box Office Average for the opening week, likely representing the average box office earnings per theater during that week.

-   **DomesticGross (Numeric)**:

    -   The total amount of money the movie earned in the domestic market (e.g., within the U.S.) during its entire run, expressed in millions of USD.

-   **ForeignGross (Numeric)**:

    -   The total box office earnings from international markets outside the domestic territory, in millions of USD.

-   **WorldGross (Numeric)**:

    -   The combined total of **DomesticGross** and **ForeignGross**, representing global earnings for the movie.

-   **Budget (Numeric)**:

    -   The total cost of producing the movie, including marketing and production expenses, in millions of USD.

-   **Profitability (Numeric)**:

    -   represent how many times the movie's global earnings (WorldGross) exceeded the production budget (Budget).

-   **OpeningWeekend (Numeric)**:

    -   The amount of money the movie made during its opening weekend, in millions of USD. This figure is often used to predict the overall success of a film.

Obtaining only the numberic values:

```{r}
movies_quant <- movies %>%
  drop_na() %>%
  select(where(is.numeric))
movies_quant
```

### Correlation between domestic gross and world gross.

```{r}
movies %>%
  gf_point(DomesticGross ~ WorldGross) %>%
  gf_lm() %>%
  gf_labs(
    title = "Scatter Plot",
    subtitle = "Movie Gross Earnings: Domestics vs World"
  )

```

There appears to be a +ve correlation between Domestic gross and world gross.

### Correlation between Profitability and Opening Week.

```{r}
movies %>%
  gf_point(Profitability ~ OpeningWeekend) %>%
  gf_lm() %>%
  gf_labs(
    title = "Scatter Plot",
    subtitle = "Movies: Does Opening Week Earnings indicate Profitability?"
  )
```

`OpeningWeek` and `Profitability` are also related in a linear way. There are just **two** movies which have been extremely profitable

### Correlation between RottenTomatoes and AudienceScore.

```{r}
movies %>%
  gf_point(RottenTomatoes ~ AudienceScore) %>%
  gf_lm() %>%
  gf_labs(
    title = "Scatter Plot",
    subtitle = "Movie Ratings: Tomatoes vs Audience"
  )
```

Rotten tomatoes and Audience score have a +ve correlation.

> We can split some of the scatter plots using one or other of the Qual variables.

### Correlation between RottenTomatoes and AudienceScore seprated by Genre.

```{r}
movies %>%
  drop_na() %>%
  gf_point(RottenTomatoes ~ AudienceScore,
    color = ~Genre
  ) %>%
  gf_lm() %>%
  gf_labs(
    title = "Scatter Plot",
    subtitle = "Movie Ratings: Trends by Genre"
  )
```

### **Quantifying Correlation**

GGally::ggpairs() : This function creates a matrix of plots for pairwise comparisons of the selected variables.

```{r}
GGally::ggpairs(
  movies %>% drop_na(),
  # Select Quant variables only for now
  columns = c(
    "RottenTomatoes", "AudienceScore", "DomesticGross", "ForeignGross"
  ),
  switch = "both",
  # axis labels in more traditional locations(left and bottom)

  progress = FALSE,
  # no compute progress messages needed

  # Choose the diagonal graphs (always single variable! Think!)
  diag = list(continuous = "barDiag"),
  # choosing histogram,not density

  # Choose lower triangle graphs, two-variable graphs
  lower = list(continuous = wrap("smooth", alpha = 0.3, se = FALSE)),
  title = "Movies Data Correlations Plot #1"
)
```

`AudienceScore` and `RottenTomatoes` are well correlated with a score of 8.3. Domestic Gross and Foreign Gross is also well correlated with a score of 8.73. Nothing else is as strongly related.

```{r}
GGally::ggpairs(
  movies %>% drop_na(),
  # Select Quant variables only for now
  columns = c(
    "Budget", "Profitability", "DomesticGross", "ForeignGross"
  ),
  switch = "both",
  # axis labels in more traditional locations(left and bottom)

  progress = FALSE,
  # no compute progress messages needed

  # Choose the diagonal graphs (always single variable! Think!)
  diag = list(continuous = "barDiag"),
  # choosing histogram,not density

  # Choose lower triangle graphs, two-variable graphs
  lower = list(continuous = wrap("smooth", alpha = 0.3, se = FALSE)),
  title = "Movies Data Correlations Plot #2"
)
```

Although not too much, you do see a correlation score of 0.65 between budget and Domestic gross and 0.67 between budget and Foreign Gross.

Budget and Profitability seem to have a slight negative correlation.

### **Doing a Correlation Test**

We try and find out how certain we can be about the correlation score by doing this.

#### What does this data tell me?

**estimate**: This is the **correlation coefficient** (often denoted as "r") between **Profitability** and **Budget** in your dataset.

**confidence level low** and **confidence level high**: These are the **lower** and **upper bounds** of uncertainity for the correlation coefficient.

**p.value**: The **p-value** shows the probability that the observed correlation happened by chance, assuming that there is no true correlation.

Movie Profitability vs Budget

```{r}
mosaic::cor_test(Profitability ~ Budget, data = movies) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 2,
    caption = "Movie Profitability vs Budget"
  )
```

Here, the correlation coefficient is -0.08 - a slight negative correlation, but we observe that the uncertainty levels lie within a large range. We also see that that p-value is high. In other words, there is **no clear linear relationship** between **Profitability** and **Budget** in your data. We, I don’t have enough evidence to make strong conclusions about a consistent negative (or positive) relationship. Increasing or decreasing a movie's budget does not reliably predict its profitability.

Movie's Domestic Gross vs Budget

```{r}
mosaic::cor_test(DomesticGross ~ Budget, data = movies) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 2,
    caption = "Movie Domestic Gross vs Budget"
  )
```

Here, the correlation coefficient is 0.7 - a positive correlation, and we observe that the uncertainty levels lie within a short range i.e. there’s low uncertainty around the estimate. We also see that that p-value is 0, concluding that this inference could not have been by chance making it statistically significant. Therefore, Budget has a strong positive effect on Profitability. Movies with higher budgets tend to have significantly higher Domestic Gross numbers, based on this data. This relationship is statistically significant, reliable, and consistent.

Movie Foreign Gross vs Budget

```{r}
mosaic::cor_test(ForeignGross ~ Budget, data = movies) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 2,
    caption = "Movie Foreign Gross vs Budget"
  )
```

Has similar inferences as the previous test.

### **The ErrorBar Plot for Correlations**
