---
title: "Impact of Visual Narratives on Public Perception"
author: "Sneha Manu Jacob"
date: "2024-11-11"
categories: [Descriptive Analysis, Statistics]
---

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(skimr)
library(ggformula)
library(crosstable)
library(infer) 
library(patchwork) 
library(ggprism)
library(supernova) 
```

## Defining the Experiment:

**Objective:** What impact does a video (any video, be it real or fake) presenting a negative narrative (even if it is clearly extremist) have on attitudes towards the entirety of a specific demographic group.

**Hypothesis:**

Null Hypothesis: The video has no impact on the viewers existing perspective.

Alternate hypothesis: A video with an extremist narrative targeting a particular demographic will negatively impact viewers' attitudes towards the entire population associated with that demographic.

**Experiment Design:**

1.  **Participants**\
    Select a diverse group of participants with varying prior attitudes towards Islam to minimize pre-existing bias. This group should ideally be balanced across demographics like age, gender, and educational background.

2.  **Pre-Exposure Opinion Score**

    -   Each participant is asked to rate their attitudes toward Islam on a scale (e.g., -10 to 10, where -10 = very negative and 10 = very positive).

3.  **Exposure to Video**

    -   Participants watch a video that presents an extremist or negative narrative about Islam.

    -   The video is designed to influence opinions through emotionally charged or polarizing content that could potentially lead to and or agree with exisiting biases or stereotypes.

4.  **Post-Exposure Opinion Score**

    -   After watching the video, participants are asked to rate their attitudes toward Islam again using the same scale.

5.  **Data Collection**

    -   Collect both pre- and post-exposure scores to quantify the change in attitudes.

Form passed around: <https://forms.gle/taP3QGedSHRc5cpx5>

### Reading and Analyzing the data:

```{r}
#| label: Read the data
visual_data <- read.csv("../../data/Visual_Responses.csv")
visual_data
```

```{r}
visual_data %>%
  select(
    Before.Exposure.to.video,
    Post.exposure.to.video
  )%>%
  pivot_longer(
    cols = everything(),
    names_to = "Time",
    values_to = "Value"
  ) -> visual_long

visual_long
```

```{r}
glimpse(visual_data)
```

```{r}
skim(visual_data)
```

### Defining the Data Dictionary:

| Variable name | Description | Type of variable |
|------------------------|------------------------|------------------------|
| ID | Assigned to each participant to keep sources anonymous | Qualitative |
| Before | The opinion score recorded before watching the video | Quantitative |
| After | The opinion score recorded before watching the video | Quantitative |

## Descriptive Analysis:

**Distribution of overall scores before:**

```{r}
visual_data %>%
  gf_histogram(~Before.Exposure.to.video)
```

Observation: This distribution is skewed in the negative direction

**Distribution of overall scores after:**

```{r}
visual_data %>%
  gf_histogram(~Post.exposure.to.video)
```

Observation: This distribution is skewed in the negative direction

```{r}
visual_long %>%
  gf_density(
    ~ Value,
    fill = ~ Time,
    alpha = 0.5,
    title = "Opinion Score Densities by Time ",
    subtitle = "Before vs After"
  )
```

```{r}
visual_new <- visual_data %>%
  mutate(Difference = Before.Exposure.to.video - Post.exposure.to.video)
visual_new
```

```{r}
visual_new %>%
  gf_density(
    ~ Difference,
    title = "Opinion Score Density difference",
    subtitle = "Before - After"
  )
```

```{r}
ggplot(visual_data, aes(x = as.factor(ID))) +
  geom_bar(aes(y = Before.Exposure.to.video, fill = "Before.Exposure.to.video"), 
           stat = "identity", position = "dodge") +
  geom_bar(aes(y = Post.exposure.to.video, fill = "Post.exposure.to.video"), 
           stat = "identity", position = "dodge", alpha = 0.6) +
  labs(x = "Participant ID", y = "Opinion Score", fill = "Condition") +
  theme_minimal() +
  scale_fill_manual(values = c("Before.Exposure.to.video" = "gold", 
                               "Post.exposure.to.video" = "blue"))
```


