---
title: "Day 2"
author: "Sneha Jacob"
date: "2024-09-27"
categories: [summaries]
---

## Introduction

Throwing away data to grasp it. I am trying to reduce data into a summarized form.

## R-code:

> note to self: don't forget to put the label!!!!!

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(skimr)
library(knitr)
library(kableExtra)
```

## The mpg data-set (car stuff):

```{r}
mpg
```

> making it look fancy. Arvind wrote this, i only added it here cause i wanted the full form of all columns cause i don't know car stuff.

```{r}
mpg %>%
  head(10) %>%
  kbl(
    # add Human Readable column names
    col.names = c(
      "Manufacturer", "Model", "Engine\nDisplacement",
      "Model\n Year", "Cylinders", "Transmission",
      "Drivetrain", "City\n Mileage", "Highway\n Mileage",
      "Fuel", "Class\nOf\nVehicle"
    ),
    caption = "MPG Dataset"
  ) %>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )



```

#### Glimpse:

```{r}
mpg %>% dplyr::glimpse()
```

> Through this glimpse, i find that cylinder is encoded as an int, and so it is considered a quantitative variable. i need to change it.

```{r}
mpg_modified <- mpg %>%
  dplyr::mutate(
    cyl = as_factor(cyl),
    fl = as_factor(fl),
    drv = as_factor(drv),
    class = as_factor(class),
    trans = as_factor(trans)
  )
glimpse(mpg_modified)
```

#### Inspecting the data-set with mosaic and skimr

A mosaic inspect of the mgp data-set only to see the difference between this and the newly created mpg_modified data set

```{r}
mpg %>% mosaic::inspect()
```

```{r}
mpg_modified %>% mosaic::inspect()
```

```{r}
mpg_modified %>% skimr::skim()
```

#### Groups within columns -

We can group a quantitative data based on different qualitative data-sets.

> Here i am finding the separate means of highway mileage for every separate level of cyl.

```{r}
mpg_modified %>%
  group_by(cyl) %>%
  summarize(average_hwy = mean(hwy), count = n())
```

> You can find mean based on 2 or more factors i.e. permutations and combinations of all factors

```{r}
mpg_modified %>%
  group_by(cyl, fl) %>%
  summarize(average_hwy = mean(hwy), count = n())
```

```{r}
mpg_modified %>%
  group_by(manufacturer) %>%
  summarize(average_cty = mean(cty))
```

## Math Anxiety Data-Set:

```{r}
math_anx <- read_delim("../../data/MathAnxiety.csv", delim=";")
math_anx
```

> AMAS (Adult Manifest Anxiety Scale), RCMAS (Revised Children's Manifest Anxiety Scale), Arith (Arithmetic Subtest of Intelligence Tests)

```{r}
math_anx %>% dplyr::glimpse()
```

```{r}
math_anx_modified <- math_anx %>%
  dplyr::mutate(
    Gender = as_factor(Gender),
    Grade = as_factor(Grade),
    Arith = as_factor(Arith)
  )
glimpse(math_anx_modified)
```

```{r}
math_anx_modified %>% skimr::skim()
```

#### Anxiety based on Gender:

```{r}
math_anx_modified %>%
  group_by(Gender) %>%
  summarize(average_AMAS = mean(AMAS), average_RCMAS = mean(RCMAS), count =n())

```

Hypothesis to Evaluate: Girls have a higher anxiety level.

#### Anxiety based on Grade:

```{r}
math_anx_modified %>%
  group_by(Grade) %>%
  summarize(average_AMAS = mean(AMAS), average_RCMAS = mean(RCMAS), count =n())

```

Hypothesis to Evaluate: Primary school students experience more anxiety than secondary school students.

#### Anxiety based on Grade and Gender:

```{r}
math_anx_modified %>%
  group_by(Grade, Gender) %>%
  summarize(average_AMAS = mean(AMAS), average_RCMAS = mean(RCMAS), count =n())
```

Hypothesis to Evaluate: Girls studying in secondary school have the highest anxiety and Secondary School boys have the least amount of anxiety.

#### Anxiety based on Arithmetic Subtest of Intelligence Tests:

```{r}
math_anx_modified %>%
  group_by(Arith) %>%
  summarize(average_AMAS = mean(AMAS), average_RCMAS = mean(RCMAS), count =n())
```

```         
```
