---
title: "Day 2"
author: "Sneha Jacob"
date: "2024-09-27"
categories: [summaries, mgp-data, math-anxiety-data]
---

## Introduction

Throwing away data to grasp it. I am trying to reduce data into a summarized form.

## R-code:

> note to self: don't forget to put the label!!!!!

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(skimr)
library(knitr)
library(kableExtra)
```

## The mpg data-set (car stuff):

```{r}
mpg
```

> making it look fancy. Arvind wrote this, i only added it here cause i wanted the full form of all columns cause i don't know car stuff.

```{r}
mpg %>%
  head(10) %>%
  kbl(
    # add Human Readable column names
    col.names = c(
      "Manufacturer", "Model", "Engine\nDisplacement",
      "Model\n Year", "Cylinders", "Transmission",
      "Drivetrain", "City\n Mileage", "Highway\n Mileage",
      "Fuel", "Class\nOf\nVehicle"
    ),
    caption = "MPG Dataset"
  ) %>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )



```

#### Glimpse:

```{r}
mpg %>% dplyr::glimpse()
```

### Table of Variables

| Variable name | Description | Type of variable |
|----|----|----|
| manufacturer | Represents the name of the company that produces the vehicle. | Qualitative |
| model | This variable indicates the specific model of the vehicle produced by the manufacturer. | Qualitative |
| displ | Engine Displacement: measures the engine's size or capacity | Quantitative |
| year | Indicates the year in which the vehicle model was manufactured. | Qualitative |
| cyl | Specifies the number of cylinders in the vehicle's engine. | Qualitative |
| trans | Indicates the type of transmission system used in the vehicle | Qualitative |
| drv | Describes how power is delivered to the wheels (e.g., front-wheel drive, rear-wheel drive, all-wheel drive) | Qualitative |
| cty | Measures the fuel efficiency of the vehicle when driving in urban conditions | Quantitative |
| hwy | Measures fuel efficiency on highways | Quantitative |
| fl | Indicates the type of fuel used by the vehicle | Qualitative |
| class | Categorizes vehicles into different classes based on their size, purpose, or design | Qualitative |

> Through this glimpse, i find that cylinder is encoded as an int, and so it is considered a quantitative variable. i need to change it.

```{r}
mpg_modified <- mpg %>%
  dplyr::mutate(
    cyl = as_factor(cyl),
    fl = as_factor(fl),
    drv = as_factor(drv),
    class = as_factor(class),
    trans = as_factor(trans)
  )
glimpse(mpg_modified)
```

#### Inspecting the data-set with mosaic and skimr

A mosaic inspect of the mgp data-set only to see the difference between this and the newly created mpg_modified data set

```{r}
mpg %>% mosaic::inspect()
```

```{r}
mpg_modified %>% mosaic::inspect()
```

```{r}
mpg_modified %>% skimr::skim()
```

### Groups within columns -

We can group a quantitative data based on different qualitative data-sets.

#### Here i am finding the separate means of highway mileage for every separate level of cyl.

```{r}
mpg_modified %>%
  group_by(cyl) %>%
  summarize(average_hwy = mean(hwy), count = n())
```

The most highway mileage is for 4 cylinders while the lowest is for 8 cylinders.

> You can find mean based on 2 or more factors i.e. permutations and combinations of all factors

#### Separate means of highway mileage for every separate level of cyl and fuel.

```{r}
mpg_modified %>%
  group_by(cyl, fl) %>%
  summarize(average_hwy = mean(hwy), count = n())
```

The highest mean highway mileage is for a cars with 4 cylinders with diesel fuel while the lowest is for electric cars with 6 cylinders.

#### Separate means of city mileage for every separate level of cyl and fuel.

```{r}
mpg_modified %>%
  group_by(cyl, fl) %>%
  summarize(average_cty = mean(cty), count = n())
```

The highest mean city mileage is for a cars with 4 cylinders with diesel fuel while the lowest is for cars with 8 cylinders and petrol fuel.

#### Separate means of city mileage for every separate level of cyl and fuel.

```{r}
mpg_modified %>%
  group_by(manufacturer) %>%
  summarize(average_cty = mean(cty))
```

The highest mean city mileage is for a cars with 4 cylinders with diesel fuel while the lowest is for electric cars with 6 cylinders.

## Math Anxiety Data-Set:

```{r}
#| label: Read the data
math_anx <- read_delim("../../data/MathAnxiety.csv", delim=";")
math_anx
```

###Glimpse the data to understand which we have to change to factors.

```{r}
math_anx %>% dplyr::glimpse()
```

### Table of Variables

| Variable name | Description | Type of variable |
|----|----|----|
| Age | Represents the age of the individual | Qualitative |
| Gender | Represents the gender of the individual(male/female) | Qualitative |
| Grade | Represents the grade they study in | Qualitative |
| AMAS | Adult Manifest Anxiety Scale | Quantitative |
| RCMAS | Revised Children's Manifest Anxiety Scale | Quantitative |
| Arith | Arithmetic Subtest of Intelligence Tests | Quantitative |

Target Variable: AMAS, RCMAS and Arith Predictor Variables: Age, Gender and Grade

### Mutate to facors:

> Among all variables, Gender, Grade and Arith must be turned into factors.

```{r}
math_anx_modified <- math_anx %>%
  dplyr::mutate(
    Gender = as_factor(Gender),
    Grade = as_factor(Grade),
    Arith = as_factor(Arith)
  )
glimpse(math_anx_modified)
```

### Summerising the data with skim:

```{r}
math_anx_modified %>% skimr::skim()
```

#### Anxiety based on Gender:

```{r}
math_anx_modified %>%
  group_by(Gender) %>%
  summarize(average_AMAS = mean(AMAS), average_RCMAS = mean(RCMAS), count =n())

```

Hypothesis to Evaluate: Girls have a higher anxiety level.

#### Anxiety based on Grade:

```{r}
math_anx_modified %>%
  group_by(Grade) %>%
  summarize(average_AMAS = mean(AMAS), average_RCMAS = mean(RCMAS), count =n())

```

Hypothesis to Evaluate: Primary school students experience more anxiety than secondary school students.

#### Anxiety based on Grade and Gender:

```{r}
math_anx_modified %>%
  group_by(Grade, Gender) %>%
  summarize(average_AMAS = mean(AMAS), average_RCMAS = mean(RCMAS), count =n())
```

Hypothesis to Evaluate: Girls studying in secondary school have the highest anxiety and Secondary School boys have the least amount of anxiety.

#### Anxiety based on Arithmetic Subtest of Intelligence Tests:

```{r}
math_anx_modified %>%
  group_by(Arith) %>%
  summarize(average_AMAS = mean(AMAS), average_RCMAS = mean(RCMAS), count =n())
```

I don't really know how to read this last data.

```         
```
