---
title: "Day 7"
author: "Sneha Manu Jacob"
date: "2024-10-18"
categories: [ Statistics]
---

## Introduction:

We are working on **Inference for Comparing Two Paired Means**

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(broom) # Tidy Test data
library(resampledata3) # Datasets from Chihara and Hesterberg's book
library(gt) # for tables
library(openintro)
```

# **Comparing Multiple Means with ANOVA**

## **Case Study #1: Results from a Diving Championship**

### **Inspecting and Charting Data**

```{r}
data("Diving2017", package = "resampledata3")
Diving2017
Diving2017_inspect <- inspect(Diving2017)
Diving2017_inspect$categorical
Diving2017_inspect$quantitative
```

```{r}
Diving2017_long <- Diving2017 %>%
  pivot_longer(
    cols = c(Final, Semifinal),
    names_to = "race",
    values_to = "scores"
  )
Diving2017_long
```

```{r}
Diving2017_long %>%
  gf_density(~scores,
    fill = ~race,
    alpha = 0.5,
    title = "Diving Scores"
  ) %>%
  gf_facet_grid(~race) %>%
  gf_fitdistr(dist = "dnorm")
```

The data are not normally distributed.

```{r}
Diving2017_long %>%
  gf_col(
    fct_reorder(Name, scores) ~ scores,
    fill = ~race,
    alpha = 0.5,
    position = "dodge",
    xlab = "Scores",
    ylab = "Name",
    title = "Diving Scores"
  )
```

-   There is no immediately identifiable **trend** in score changes from one race to the other.

```{r}
Diving2017_long %>%
  gf_boxplot(
    scores ~ race,
    fill = ~race,
    alpha = 0.5,
    xlab = "Race",
    ylab = "Scores",
    title = "Diving Scores"
  )
```

-   Although the two *medians* appear to be different, the box plots overlap considerably. So one cannot visually conclude that the two sets of race timings have different *means*.

# **Comparing Multiple Means with ANOVA**

## Frog data: 

```{r}
library(patchwork) # Arranging Plots
library(ggprism) # Interesting Categorical Axes
library(supernova) # Beginner-Friendly ANOVA Tables
```

```{r}
frogs_orig <- read_csv("../../data/frogs.csv")
frogs_orig
```

Our response variable is the hatching `Time`. Our explanatory variable is a *factor*, `Temperature`, with 3 levels: 13°C, 18°C and 25°C. 

### Clean the data:

```{r}
frogs_orig %>%
  pivot_longer(
    .,
    cols = starts_with("Temperature"),
    cols_vary = "fastest",
    # new in pivot_longer
    names_to = "Temp",
    values_to = "Time"
  ) %>%
  drop_na() %>%
  ##
  separate_wider_regex(
    cols = Temp,
    # knock off the unnecessary "Temperature" word
    # Just keep the digits thereafter
    patterns = c("Temperature", TempFac = "\\d+"),
    cols_remove = TRUE
  ) %>%
  # Convert Temp into TempFac, a 3-level factor
  mutate(TempFac = factor(
    x = TempFac,
    levels = c(13, 18, 25),
    labels = c("13", "18", "25")
  )) %>%
  rename("Id" = `Frogspawn sample id`) -> frogs_long
frogs_long
##
frogs_long %>% count(TempFac)
```

1.  The dataset is reshaped from wide to long format, collecting temperature data into a single `Temp` column.

2.  Missing data is dropped.

3.  The `Temp` column is cleaned to retain only the numeric values (13, 18, 25).

4.  The temperature values are converted into a factor with three levels.

5.  The sample ID column is renamed.

6.  Finally, a count is performed to see how many observations fall into each of the temperature categories (13, 18, or 25°C).

This process cleans and organizes the data, making it easier to analyze the effect of temperature on the frogs' data.

### Data Visualizations:

```{r}
gf_histogram(~Time,
  fill = ~TempFac,
  data = frogs_long, alpha = 0.5
) %>%
  gf_vline(xintercept = ~ mean(Time)) %>%
  gf_labs(
    title = "Histograms of Hatching Time Distributions vs Temperature",
    x = "Hatching Time", y = "Count"
  ) %>%
  gf_text(7 ~ (mean(Time) + 2),
    label = "Overall Mean"
  ) %>%
  gf_refine(guides(fill = guide_legend(title = "Temperature level (°C)")))
```

Temprature and hatching time seem to have a +ve corelation.

```{r}
gf_boxplot(
  data = frogs_long,
  Time ~ TempFac,
  fill = ~TempFac,
  alpha = 0.5
) %>%
  gf_vline(xintercept = ~ mean(Time)) %>%
  gf_labs(
    title = "Boxplots of Hatching Time Distributions vs Temperature",
    x = "Temperature", y = "Hatching Time",
    caption = "Using ggprism"
  ) %>%
  gf_refine(
    scale_x_discrete(guide = "prism_bracket"),
    guides(fill = guide_legend(title = "Temperature level (°C)"))
  )
```

Temperature seems to have a huge effect of hatching time

## **ANOVA**

A statistical test for 3 or more means:

```{r}
frogs_anova <- aov(Time ~ TempFac, data = frogs_long)
```

ErrorBar plot

```{r}
supernova::pairwise(frogs_anova,
  correction = "Bonferroni", # Try "Tukey"
  alpha = 0.05, # 95% CI calculation
  var_equal = TRUE, # We'll see
  plot = TRUE
)
```

```{r}
supernova::pairwise(frogs_anova,
  correction = "Tukey",
  alpha = 0.05, # 95% CI calculation
  var_equal = TRUE, # We'll see
  plot = TRUE
)
```

The black points: Estimate difference in means of the hatching time i.e. differences in mean pair vise.

## Proportions:

CLT for proportions
 1. sample proportion = population propotion
 2. only go ahead with a sample if it has atleast 10 of the "minority"
 
 
```{r}
data(yrbss, package = "openintro")
yrbss
```

